---
title: ""
output: html_document
---

# Dataset 1: Peruvian Blood Pressure

This dataset consists of variables possibly relating to blood pressures of 39 Peruvians who have moved from rural high-altitude areas to urban lower-altitude areas. The variables in this dataset are listed as follows: 

- `Age`: Age in years
- `Years`: Years in urban area
- `Weight`: Weight in kg
- `Height`: Height in mm
- `Chin`: Chin skinfold
- `Forearm`: Forearm skinfold
- `Calf`: Calf skinfold
- `Pulse`: Resting pulse rate
- `Systol`: Systolic blood pressure

You work in a pharmaceutical company and want to conduct suitable statistical analysis to understand these data. In particular, you are asked to find significant variables that contribute to high systolic blood pressure. 

The data are in the `idarps` R package. Make sure to update it using:

```{r, eval=FALSE}
install.packages("idarps")
```

Then, you can load this dataset as follows:

```{r, eval=FALSE}
library(idarps)
data("PeruvianBP")
head(PeruvianBP)
```


# Dataset 2: Breast Cancer 

This dataset consists of several clinical features observed or measured for 116 participants in a study of breast cancer. In particular, age, BMI, glucose, insulin and HOMA are recorded for each participant, as well as whether they have breast cancer or not. 

- `Age`: Age in years
- `BMI`: Body mass index in kg/$m^2$
- `Glucose`: Glucose in mg/dL
- `Insulin`: Insulin in $\mu$U/mL
- `HOMA`: Homeostasis model assessment 
- `Classification`: Presence of breast cancer (0 if no cancer, 1 if with cancer)


You work in a hospital and your goal is to develop a suitable model based on the available data in order to discover the significant features that contribute to the presence of breast cancer in women. 

The data are in the `idarps` R package and you can load the data as follows:

```{r, eval=FALSE}
library(idarps)
data("BreastCancer")
head(BreastCancer)
```


