---
title: "Homework 2"
output: html_document
---

In this homework, we wish to investigate Abouir, et al., (2022) to assess the validity of several claims. First of all, we need to select the observations for different analysis.

```{r}
library(idar)
data(codex)

# Cmax
Cmax_men = codex$cmax[codex$gender == 0]
Cmax_women = codex$cmax[codex$gender == 1]
Cmax_non_obese = codex$cmax[codex$obese == 0]
Cmax_obese = codex$cmax[codex$obese == 1]

# Tmax
Tmax_men = codex$tmax[codex$gender == 0]
Tmax_women = codex$tmax[codex$gender == 1]
Tmax_non_obese = codex$tmax[codex$obese == 0]
Tmax_obese = codex$tmax[codex$obese == 1]

# AUC
AUC_men = codex$auc[codex$gender == 0]
AUC_women = codex$auc[codex$gender == 1]
AUC_non_obese = codex$auc[codex$obese == 0]
AUC_obese = codex$auc[codex$obese == 1]
```

### Question 1: Difference between groups for $C_{\max}$

Let's have a look at the empirical distributions:

```{r, fig.align='center', fig.height=6, fig.width=12}
par(mfrow = c(1, 2))
boxplot_w_points(Cmax_women, Cmax_men, names = c("Women", "Men"), ylab = "Cmax (ng/ml)")
boxplot_w_points(Cmax_obese, Cmax_non_obese, names = c("Obese", "Non-obese"), ylab = "Cmax (ng/ml)")
```

Based on these graphs, it is reasonable to believe that some outliers are present. Thus, we select Wilcoxon test to compare the means of the different groups.

```{r, warning=F}
wilcox.test(Cmax_men, Cmax_women, alternative = "less")
wilcox.test(Cmax_obese, Cmax_non_obese, alternative = "less")
```

From the result we can see that the p-value = 0.001661 in the first test is smaller than 0.05, which indicates there are some evidence against the null hypotheses in favor of the alternative. Therefore, at the 95% confidence level, we reject the null and accept the alternative hypothesis that the mean of the variable `cmax` in group of `women` is smaller than that of the group `men`. However, for the second test, there isn't enough an evidence to reject the null hypothesis which we assert that the means for the variable `cmax` are the same between the groups `obese` and `non_obese`. Interestingly, different conclusions would be obtained if we were use a (Welch) t-test:

```{r}
t.test(Cmax_men, Cmax_women, alternative = "less")
t.test(Cmax_obese, Cmax_non_obese, alternative = "less")
```

Indeed, we obtain a p-value smaller than $\alpha = 0.05$ for the second test leading to a different conclusion.

### Question 2: Difference between groups for $T_{\max}$

We proceed similarly for the second question:

```{r, fig.align='center', fig.height=6, fig.width=12}
par(mfrow = c(1, 2))
boxplot_w_points(Tmax_women, Tmax_men, names = c("Women", "Men"), ylab = "Tmax (h)")
boxplot_w_points(Tmax_obese, Tmax_non_obese, names = c("Obese", "Non-obese"), ylab = "Tmax (h)")
```

Based on these graphs, we also select Wilcoxon test and obtain the p-values as follows:

```{r}
wilcox.test(Tmax_women, Tmax_men, alternative = "less")
wilcox.test(Tmax_obese, Tmax_non_obese, alternative = "less")
```

Thus, we fail the reject the null hypothesis at the 95% confidence level for both tests.

# ### Question 3: Difference between groups for AUC

We proceed similarly for the last question:

```{r, fig.align='center', fig.height=6, fig.width=12}
par(mfrow = c(1, 2))
boxplot_w_points(AUC_women, AUC_men, names = c("Women", "Men"), ylab = "AUC (ng.h/ml)")
boxplot_w_points(AUC_obese, AUC_non_obese, names = c("Obese", "Non-obese"), ylab = "AUC (ng.h/ml)")
```

We obtain the p-values are follows:

```{r}
wilcox.test(AUC_men, AUC_women, alternative = "less")
wilcox.test(AUC_obese, AUC_non_obese, alternative = "less")
```

Therefore, at the 95% confidence level, we reject the null hypothesis for the test and accept the alternative hypothesis stating that the mean of the variable `auc` in group of `women` is smaller than that of the group `men`. However, for the second test, there isn't enough an evidence (at the 95% confidence level) to reject the null hypothesis which we assert that the means for the variable `auc` are the same between the groups `obese` and `non_obese`.

