---
title: "Optional Homework 3: Solution"
output: html_document
---

We would like to fit the following model: 

$$
\log(\text{signal}_i) = \beta_0 + \beta_1\text{group}_i + \beta_2\text{dose}_i + \beta_3\text{group}_i\text{dose}_i + \epsilon_i
$$

```{r}
library(idarps)
data("HP13Cbicarbonate")

# Log transform
HP13Cbicarbonate$log_signal = log(HP13Cbicarbonate$signal)

# Fit the model
mod = lm(log_signal ~ group*dose, data = HP13Cbicarbonate)
summary(mod)
```

Next we evaluate graphically the adequacy of this model with the data:

```{r, fig.align='center'}
colors = c("#F8766D", "#00BFC4")
plot(HP13Cbicarbonate$dose, HP13Cbicarbonate$log_signal, 
     xlab = "Dose", ylab = "Signal (in log scale)", 
     col = colors[as.numeric(HP13Cbicarbonate$group == "Fed") + 1], 
     pch = 15 + as.numeric(HP13Cbicarbonate$group == "Fed"))
grid()
legend("topright", c("Fasted", "Fed"), col = colors, pch = c(15, 16), bty = "n")

dose_to_predict = seq(0.01, 0.3, length.out = 20)
pred_fasted = predict(mod, data.frame(dose = dose_to_predict, group = rep("Fasted", length(dose_to_predict))), interval = "confidence")
pred_fed = predict(mod, data.frame(dose = dose_to_predict, group = rep("Fed", length(dose_to_predict))), interval = "confidence")

lines(dose_to_predict, pred_fasted[,1], col = "#F8766D")
polygon(c(dose_to_predict, rev(dose_to_predict)), c(pred_fasted[,2], rev(pred_fasted[,3])), col = "#F8766D33", border = NA)

lines(dose_to_predict, pred_fed[,1], col = "#00BFC4")
polygon(c(dose_to_predict, rev(dose_to_predict)), c(pred_fed[,2], rev(pred_fed[,3])), col = "#00BFC433", border = NA)
```

We also construct the model diagnostic graphs for this model as follows:

```{r, fig.align='center'}
plot(predict(mod), resid(mod), 
     xlab = "Fitted values", ylab = "Residuals", 
     col = colors[as.numeric(HP13Cbicarbonate$group == "Fed") + 1], 
       pch = 15 + as.numeric(HP13Cbicarbonate$group == "Fed"))
```

```{r, fig.align='center'}
qqnorm(resid(mod), 
       col = colors[as.numeric(HP13Cbicarbonate$group == "Fed") + 1], 
       pch = 15 + as.numeric(HP13Cbicarbonate$group == "Fed"))
qqline(resid(mod))
```

As we can see, this model appears to be adequate to describe our data based on these diagnostic graphs.

Now we evaluate the claims of the scientists. We recall the model summary as below: 

```{r}
summary(mod)
```

**Claim 1**: "*In the fasted group, the variable dose has no impact on the signal*", which corresponds to the hypotheses:

$$H_0: \;\; \beta_2 = 0, \;\;\; H_a: \;\; \beta_2 \neq 0.$$
The p-value of this test is 0.88846. Therefore, we can conclude at the 95% confidence level that the variable dose has no impact on the signal in the fasted group.

**Claim 2**: "*In the fed group, the variable dose has an impact on the signal (and we expect that the higher is the dose, the lower is the signal)*", which corresponds to the hypotheses:

$$H_0: \;\; \beta_3 = 0, \;\;\; H_a: \;\; \beta_3 < 0.$$

The p-value of this test is 0.02199. Therefore, at the 95% confidence level we can reject the null hypothesis and conclude that in the fed group, the higher is the dose, the lower is the signal. 